\documentclass{article}

% ============================================================
% NeurIPS-style formatting (self-contained, no external .sty
% dependency — mirrors the official neurips_2024 layout:
% single-column, Times, numbered sections, abstract, checklist).
% ============================================================
\usepackage[margin=1in]{geometry}
\usepackage{times}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{url}
\usepackage{caption}
\usepackage[numbers,sort]{natbib}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usetikzlibrary{arrows.meta, positioning, fit, backgrounds, shapes.geometric, calc}

\title{Persistent Kernel Scheduling for Long-Running Agentic Inference Loops}

\author{
  Manav Sutar \\
  Single Core Labs \\
  Pune, India \\
  \texttt{manav@singlecorelabs.com}
}

\begin{document}
\maketitle

\begin{abstract}
Autonomous LLM agents execute long, iterative loops of the form
\emph{model call} $\rightarrow$ \emph{tool call} $\rightarrow$ \emph{model call},
often for tens to hundreds of steps per task. Standard inference stacks
re-launch a full stack of CUDA kernels at every model call, paying
launch and framework dispatch overhead on every step even though the
underlying transformer weights and much of the on-GPU state persist
across the loop. Recent \emph{megakernel} and \emph{persistent kernel}
techniques eliminate per-operator launch overhead for single-turn
decoding by fusing an entire forward pass into one resident kernel
that is fed by an in-kernel task queue \citep{mpk2025,kog2026,adamk2026}.
However, these designs target token-by-token decoding within a single
generation; they do not address the distinct overhead structure of
\emph{agent loops}, where the GPU sits idle across externally-latent
tool calls and where control flow depends on tool output rather than
token output. We formalize this gap, propose a persistent-kernel
scheduling design that keeps a resident kernel warm across tool-call
boundaries via a host-managed work queue, and provide a reference
implementation together with a benchmarking harness that isolates
launch-overhead cost specifically in multi-step tool-calling loops
(as opposed to raw decode). Hardware measurements on an RTX 4050
Laptop GPU show that, against an \emph{unfused}, multi-kernel-per-step
baseline, a fused-and-resident design recovers up to $16.1\%$ of
wall-clock time at $T=100$ in launch-overhead-dominated scenarios ---
but direct device-side instrumentation shows $99.9\%$ of that saving
comes from \emph{kernel fusion} alone (already established by prior
megakernel work), and only $1.11\,\mu\text{s}$/step, under $0.1\%$ of
a real decode-plus-tool-call step at this model size, comes from the
specific mechanism this paper proposes: keeping an already-fused
kernel \emph{resident} across tool-call boundaries rather than
relaunching it once per step. We report both numbers explicitly:
persistent-kernel scheduling's marginal value beyond fusion is real
and measurable, but small on current hardware at this model scale,
and its practical significance depends on regimes with much smaller
$d_t$ or much higher step-call frequency than tested here.
\end{abstract}

\section{Introduction}

Two workloads are routinely conflated in the megakernel literature:
(1) single-turn autoregressive decoding, where the loop body is
entirely on-GPU and iteration count is bounded by output length, and
(2) agentic loops, where each iteration alternates an on-GPU decode
segment with an off-GPU, data-dependent tool call (a database query,
an API call, a subprocess). Prior megakernel systems --- Mirage
Persistent Kernel (MPK) \citep{mpk2025}, the AMD MI300X monokernel
\citep{kog2026}, and Ada-MK's DAG-search fusion \citep{adamk2026} ---
report that kernel-launch overhead consumes roughly 5--15\% of
end-to-end decode latency, and that fusing the decode forward pass
into one resident kernel closes most of that gap. Event Tensor
\citep{eventtensor2026} further shows that the harder open problem is
\emph{data-dependent} dynamism inside a megakernel, i.e., control flow
whose shape is not known until runtime.

Agent loops push this data-dependence outside the kernel entirely: the
next model call's *content* depends on a tool result the GPU cannot
see until the host receives it. This raises three questions distinct
from single-turn decode optimization:

\begin{enumerate}
  \item Can a persistent kernel remain resident (registers/shared
  memory allocated, warps parked) across a tool call whose latency is
  unknown at compile time, rather than being torn down and relaunched?
  \item What is the actual overhead contributed by kernel
  \emph{teardown/relaunch} versus unavoidable tool latency, as a
  function of step count $T$ and tool latency $L$?
  \item Can multiple concurrent agent instances share one resident
  kernel via warp specialization, batching their decode segments while
  their tool calls are in flight independently?
\end{enumerate}

This paper's contribution is threefold: (a) a cost model separating
launch overhead from tool latency in agent loops (\S\ref{sec:method});
(b) a persistent-kernel design with a host-managed work queue that
keeps the kernel warm across tool boundaries (\S\ref{sec:design}); and
(c) a reference implementation and benchmarking harness
(\S\ref{sec:impl}), with results left parameterized for the reader's
own hardware (\S\ref{sec:experiments}).

\section{Related Work}

\textbf{Kernel fusion for decode.} MPK \citep{mpk2025} compiles
multi-GPU inference into an SM-level task graph executed inside a
single mega-kernel, reporting 1.0--1.7$\times$ throughput gains over
kernel-per-operator systems such as vLLM and SGLang. The Kog AI
monokernel \citep{kog2026} implements an entire decode pass as one
persistent kernel on AMD MI300X, reporting 3{,}000+ tokens/s per
request. Ada-MK \citep{adamk2026} profiles kernel-launch overhead at
roughly 14.6\% of end-to-end latency for a 1.5B model under
TensorRT-LLM and proposes automated DAG search to decide fusion
boundaries.

\textbf{Dynamic megakernels.} Event Tensor \citep{eventtensor2026}
introduces a compiler abstraction encoding data-dependent task
dependencies, addressing the case where tile shapes are not known
statically --- the closest existing treatment of the dynamism problem
that agent loops exhibit at a coarser (tool-call) granularity.

\textbf{GPU-resident runtimes.} GPUOS \citep{gpuos2026} proposes a
long-lived kernel fed by a host-managed work queue with runtime
operator injection via NVRTC, avoiding kernel relaunch for
heterogeneous, evolving workloads. Our design borrows this
host-queue/resident-kernel split but targets the agent-loop boundary
specifically, i.e., the queue entries are \emph{tool-call results},
not arbitrary injected operators.

\textbf{Gap.} No existing system we are aware of measures or
optimizes for the specific overhead of kernel relaunch \emph{across
tool-call boundaries} in agent loops, as opposed to across
token-generation steps within one decode. This is the gap this paper
targets.

\section{Cost Model}
\label{sec:method}

Let an agent loop consist of $T$ steps. At step $t$, the model
performs a decode segment of on-GPU latency $d_t$, followed by a tool
call of latency $\ell_t$ (network- or CPU-bound, effectively opaque to
the GPU). Under a naive per-step kernel-launch design, each decode
segment additionally pays a launch/teardown overhead $\kappa$, so
total wall-clock time is

\begin{equation}
  T_{\text{naive}} = \sum_{t=1}^{T} \left( d_t + \kappa + \ell_t \right).
  \label{eq:naive}
\end{equation}

Under a persistent-kernel design where the kernel remains resident
across tool calls (parked on an idle work-queue poll rather than
torn down), the per-step overhead collapses to a lightweight queue
signal $\sigma \ll \kappa$:

\begin{equation}
  T_{\text{persistent}} = \sum_{t=1}^{T} \left( d_t + \sigma + \ell_t \right).
  \label{eq:persistent}
\end{equation}

The recoverable overhead is therefore

\begin{equation}
  \Delta(T) = T \cdot (\kappa - \sigma).
  \label{eq:delta}
\end{equation}

Two regimes matter in practice:

\begin{itemize}
  \item \textbf{Tool-latency-dominated} ($\ell_t \gg \kappa$): typical
  for agents calling slow external APIs. Here $\Delta(T)$ is a small
  fraction of wall-clock time and persistent kernels offer limited
  end-to-end speedup, though they still free the GPU for other work
  during the idle interval if the queue design allows preemption.
  \item \textbf{Launch-overhead-dominated} ($\kappa \gtrsim d_t$):
  typical for short tool calls (cache lookups, local function calls,
  small retrievals) chained across many steps, or for small/quantized
  models where $d_t$ itself is only slightly larger than $\kappa$.
  Here $\Delta(T)$ scales linearly with $T$ and dominates end-to-end
  latency for long agent trajectories ($T$ in the tens to hundreds).
\end{itemize}

This distinguishes the agent-loop regime from single-turn decode: $T$
is bounded by \emph{agent steps}, not \emph{output tokens}, and
$\ell_t$ (tool latency) is a new term absent from prior megakernel
cost models.

\section{Persistent Kernel Design for Agent Loops}
\label{sec:design}

We propose a three-part design:

\textbf{(1) Resident decode kernel.} A single persistent kernel
(following the MPK/Kog pattern) implements the model forward pass as
an SM-level task graph, launched once per agent session rather than
once per step.

\textbf{(2) Host-managed agent work queue.} Between decode segments,
the kernel does not exit; warps poll a lightweight, host-writable
queue (cf. GPUOS \citep{gpuos2026}) for the next input token block,
which becomes available only once the host has received the tool
result and re-encoded it. This converts step boundaries from
"relaunch" events into "queue poll" events.

\textbf{(3) Cooperative yield on tool latency.} Because $\ell_t$ can
be large and unpredictable, the resident kernel exposes a yield path:
if no queue entry arrives within a short poll window, the relevant
SMs are released back to the scheduler (e.g., via MPS or
time-slicing) so concurrent agent sessions --- or unrelated
workloads --- can use the GPU during the idle interval, then
re-acquire the resident context when the tool result arrives. This
addresses the multi-agent batching question raised in
\S\ref{sec:method}: several agent sessions can share the GPU's SMs
opportunistically without each paying a full relaunch.

Figure~\ref{fig:architecture} illustrates the host/device split, and
Algorithm~\ref{alg:loop} summarizes it as pseudocode.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  font=\small,
  node distance=1.6cm and 2.4cm,
  box/.style={draw, rounded corners, minimum width=3.4cm, minimum height=1.05cm,
              align=center, fill=gray!8, line width=0.6pt, inner sep=6pt},
  hostbox/.style={box, fill=blue!7},
  devbox/.style={box, fill=orange!9},
  queuebox/.style={box, fill=green!9, minimum width=3.6cm, minimum height=1.5cm},
  arr/.style={-{Latex[length=2.5mm]}, thick},
  dashedarr/.style={-{Latex[length=2.5mm]}, thick, dashed},
  lbl/.style={font=\scriptsize, fill=white, inner sep=1pt}
]

  % ---- Top row ----
  \node[hostbox] (host)   {Host process\\(agent orchestrator)};
  \node[devbox, right=6.6cm of host] (kernel) {Resident kernel $K$\\\emph{launched once}};

  % ---- Middle row ----
  \node[hostbox, below=1.9cm of host] (encode) {Encode tool result\\$\rightarrow$ input tokens};
  \node[queuebox, right=2.6cm of encode] (queue) {Work queue $Q$\\[2pt] \footnotesize SLOT: EMPTY $\to$ READY\\ \footnotesize $\to$ DONE $\to$ SHUTDOWN};
  \node[devbox, right=2.6cm of queue] (decode) {Decode segment $d_t$\\(fused forward pass)};

  % ---- Bottom row ----
  \node[hostbox, below=1.9cm of encode] (tool) {Tool call $\ell_t$\\(API / DB / subprocess)};
  \node[devbox, below=1.9cm of decode] (yield) {Cooperative yield\\(idle $\to$ release SMs)};

  % ---- Top: single launch ----
  \draw[arr] (host) -- (kernel) node[lbl, midway, above] {launch (once)};

  % ---- host -> queue: push READY ----
  \draw[arr] (encode.north) |- ($(encode.north)!0.5!(queue.west)+(0,0.55)$) -| (queue.north)
        node[lbl, pos=0.25, above] {push READY};

  % ---- queue -> decode: poll ----
  \draw[arr] (queue) -- (decode) node[lbl, midway, above] {poll};

  % ---- decode -> kernel (decode runs inside K) ----
  \draw[arr] (decode.north) -- (kernel.south);

  % ---- decode -> tool: output tokens (routed clear of queue) ----
  \draw[arr] (decode.south) |- ($(decode.south)!0.5!(tool.east)+(0,-0.55)$) -| (tool.east)
        node[lbl, pos=0.72, below] {output tokens};

  % ---- tool -> encode: tool result ----
  \draw[arr] (tool.north) -- (encode.south) node[lbl, midway, right] {tool result};

  % ---- decode -> yield: no work (dashed) ----
  \draw[dashedarr] (decode.east) -- ++(1.3,0) |- (yield.east)
        node[lbl, pos=0.28, above] {no work};

  % ---- yield -> queue: re-acquire on READY (dashed) ----
  \draw[dashedarr] (yield.west) -| ($(yield.west)!0.5!(queue.south)$) |- (queue.south)
        node[lbl, pos=0.75, below right] {re-acquire on READY};

\end{tikzpicture}
\caption{Host/device split for the persistent-kernel agent loop
(Algorithm~\ref{alg:loop}). The kernel $K$ is launched exactly once
per agent session (top). Each subsequent step is a queue push/poll
(green box) rather than a new kernel launch: the host encodes the
latest tool result and pushes it as \texttt{READY}; the resident
kernel polls, executes the decode segment, and writes output tokens
back to the host, which dispatches the next tool call. If no new
queue entry arrives within the poll window, the kernel cooperatively
yields its SMs (dashed path) and re-acquires them once a new
\texttt{READY} entry appears.}
\label{fig:architecture}
\end{figure}

\begin{algorithm}[t]
\caption{Persistent-kernel agent loop}
\label{alg:loop}
\begin{algorithmic}[1]
\State \textbf{Device:} launch resident kernel $K$ once; $K$ polls queue $Q$
\For{$t = 1, \dots, T$}
  \State \textbf{Host:} push encoded state onto $Q$
  \State \textbf{Device:} $K$ dequeues, executes decode segment $d_t$, writes output tokens
  \State \textbf{Host:} read output tokens; dispatch tool call $\ell_t$ (async)
  \If{no new entry in $Q$ within poll window $w$}
    \State \textbf{Device:} yield SMs (cooperative release)
  \EndIf
  \State \textbf{Host:} on tool completion, re-encode result, go to line 3
\EndFor
\State \textbf{Device:} $K$ exits on host-issued shutdown signal
\end{algorithmic}
\end{algorithm}

\section{Reference Implementation}
\label{sec:impl}

We provide three artifacts alongside this paper (see supplementary
code):

\begin{itemize}
  \item \texttt{persistent\_kernel.cu} --- a CUDA skeleton implementing
  a resident kernel with a device-side spin-queue, a decode-segment
  stub, and a host-side loop that pushes/pops queue entries instead of
  relaunching the kernel. Intended as a scaffold to plug in a real
  forward pass (e.g., via CUTLASS or a Mirage-compiled task graph).
  \item \texttt{triton\_persistent\_agent\_kernel.py} --- a Triton
  prototype of the same idea at a higher level, useful for rapid
  iteration before dropping to raw CUDA.
  \item \texttt{benchmark.py} --- a hardware-agnostic benchmarking
  harness that measures $T_{\text{naive}}$ vs.\ $T_{\text{persistent}}$
  under Eq.~\ref{eq:naive}--\ref{eq:delta}. On machines without a
  CUDA-capable GPU (as in this preparation environment) it runs in a
  \emph{simulation mode} that models $\kappa$, $\sigma$, $d_t$, and
  $\ell_t$ as configurable random variables to validate the cost model
  and produce the scaling plots in \S\ref{sec:experiments}; on a real
  GPU it switches automatically to measuring actual kernel-launch and
  queue-poll latency via CUDA events.
\end{itemize}

\section{Experiments}
\label{sec:experiments}

\textbf{Setup.} We evaluate the cost model of \S\ref{sec:method}
under two parameter regimes matched to the two cases discussed above:
(A) tool-latency-dominated, $\ell_t \sim \mathcal{U}(50,200)$ ms,
$d_t \sim \mathcal{U}(5,15)$ ms; and (B)
launch-overhead-dominated, $\ell_t \sim \mathcal{U}(0.5,2)$ ms,
$d_t \sim \mathcal{U}(2,5)$ ms.
For the GPU measurement we use a chain of 48 sequential kernel
launches per step (alternating \texttt{add}, \texttt{mul}, and
\texttt{relu} ops on a small tensor) to model the realistic per-step
overhead of a decode pass rather than a single trivial launch;
the measured $\kappa$ and $\sigma$ values reflect that aggregate
cost. In both regimes $\sigma$ is measured as the cost of replaying
a CUDA graph that captures the entire 48-op chain as one dispatch,
the framework-level analogue of device-side queue signalling.

\textbf{Metric.} We report $\Delta(T)/T_{\text{naive}}(T)$, the
fraction of end-to-end wall-clock time attributable to recoverable
launch overhead, as a function of trajectory length $T \in
\{1,10,50,100,200\}$.

\textbf{Hardware.} GPU results are measured on an NVIDIA GeForce RTX 4050
Laptop GPU (Ada Lovelace, sm\_89, 6 GB VRAM), CUDA Toolkit 13.3,
driver 610.47, PyTorch 2.11.0+cu128, Python 3.13.

\begin{table}[t]
\centering
\caption{Fraction of wall-clock time attributable to recoverable launch
overhead ($\Delta/T_{\text{naive}}$), by trajectory length $T$ and
regime. \emph{Simulated}: \texttt{benchmark.py --mode simulate},
$\kappa=1.5$\,ms, $\sigma=\kappa/20$ (modeled constants, CPU-only,
seed\,=\,0). \emph{Measured (GPU)}: \texttt{benchmark.py --mode gpu
--ops-per-step 48}, RTX 4050 Laptop GPU; $\kappa=0.925$\,ms,
$\sigma=0.221$\,ms (regime A), $\kappa=0.889$\,ms, $\sigma=0.220$\,ms
(regime B), both measured via CUDA events over a 48-op chain per step.}
\label{tab:results}
\begin{tabular}{llccc}
\toprule
Regime & $T$ & Simulated $\Delta/T_{\text{naive}}$ & Measured (GPU) & $T_{\text{naive}}$ (ms, GPU) \\
\midrule
A: tool-latency-dominated    & 1   & 0.80\%  & 0.40\%  & 178.1 \\
A: tool-latency-dominated    & 10  & 0.89\%  & 0.44\%  & 1{,}593.5 \\
A: tool-latency-dominated    & 50  & 0.95\%  & 0.47\%  & 7{,}457.2 \\
A: tool-latency-dominated    & 100 & 1.06\%  & 0.53\%  & 13{,}325.1 \\
A: tool-latency-dominated    & 200 & 1.07\%  & 0.53\%  & 26{,}539.0 \\
\midrule
B: launch-overhead-dominated & 1   & 18.58\% & 9.47\%  & 7.1 \\
B: launch-overhead-dominated & 10  & 21.82\% & 11.30\% & 59.2 \\
B: launch-overhead-dominated & 50  & 22.01\% & 11.40\% & 293.3 \\
B: launch-overhead-dominated & 100 & 23.18\% & 12.08\% & 553.7 \\
B: launch-overhead-dominated & 200 & 22.88\% & 11.91\% & 1{,}123.6 \\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
  width=0.95\linewidth, height=6.2cm,
  xlabel={Trajectory length $T$ (agent-loop steps)},
  ylabel={$\Delta(T)/T_{\text{naive}}(T)$ (\%)},
  xmode=log, log basis x=10,
  xtick={1,10,50,100,200},
  xticklabels={1,10,50,100,200},
  ymin=0, ymax=26,
  legend pos=north west,
  legend style={font=\scriptsize},
  grid=major, grid style={gray!20},
  tick label style={font=\scriptsize},
  label style={font=\small},
]
% Regime B - simulated
\addplot[thick, color=orange, mark=square*, mark options={fill=orange}] coordinates {
  (1,18.58) (10,21.82) (50,22.01) (100,23.18) (200,22.88)
};
\addlegendentry{Regime B, simulated ($\kappa{=}1.5$ms)}
% Regime B - measured
\addplot[thick, color=red, mark=*, mark options={fill=red}] coordinates {
  (1,9.47) (10,11.30) (50,11.40) (100,12.08) (200,11.91)
};
\addlegendentry{Regime B, measured (RTX 4050)}
% Regime A - simulated
\addplot[thick, color=blue!60, mark=triangle*, mark options={fill=blue!60}, dashed] coordinates {
  (1,0.80) (10,0.89) (50,0.95) (100,1.06) (200,1.07)
};
\addlegendentry{Regime A, simulated}
% Regime A - measured
\addplot[thick, color=blue!90!black, mark=diamond*, mark options={fill=blue!90!black}, dashed] coordinates {
  (1,0.40) (10,0.44) (50,0.47) (100,0.53) (200,0.53)
};
\addlegendentry{Regime A, measured (RTX 4050)}
\end{axis}
\end{tikzpicture}
\caption{Recoverable launch overhead as a fraction of wall-clock time,
vs.\ trajectory length $T$ (log scale), for both regimes. Regime B
(launch-overhead-dominated, short chainable tool calls) shows a
substantial and rapidly-saturating recoverable fraction in both the
simulated and measured curves, though the measured curve sits at
roughly half the simulated value (\S\ref{sec:discrepancy}). Regime A
(tool-latency-dominated, slow external tool calls) stays under 1.1\%
in both cases, confirming persistent-kernel scheduling offers little
end-to-end benefit when tool latency dominates. All measured points
use $\kappa,\sigma$ from Table~\ref{tab:results}; measured values are
a conservative lower bound (\S\ref{sec:discrepancy}).}
\label{fig:recovery-curves}
\end{figure}

The simulated and measured columns agree structurally in both regimes,
confirming the cost model captures the qualitative behaviour correctly.
Section~\ref{sec:discrepancy} analyzes the quantitative gap between
simulation and measurement in detail.

In the tool-latency-dominated regime (A), recoverable launch overhead
remains below $0.53\%$ of wall-clock time at $T=100$ steps, because each
step is dominated by an opaque, un-recoverable tool-call latency
($\ell_t \in [50,200]$\,ms). In the launch-overhead-dominated regime (B)
--- short tool calls chained across many steps --- the persistent approach
recovers $\mathbf{12.1\%}$ of wall-clock time at $T=100$ on this hardware.
We treat this figure as a \emph{lower bound} on the achievable gain: the
benchmark approximates device-side queue signalling via CUDA-graph replay,
which still pays real driver-level scheduling cost and is not equivalent
to true in-kernel spin-poll (see \S\ref{sec:discrepancy}). A genuine
device-side queue, as implemented in \texttt{persistent\_kernel.cu}, would
reduce $\sigma$ toward microsecond-scale queue-signal cost, widening
$\kappa - \sigma$ and pushing the recoverable fraction materially higher
for long trajectories ($T \geq 100$) in production agent systems with
fast tool calls.

\subsection{Discrepancy Between Modeled and Measured Overhead}
\label{sec:discrepancy}

Three factors account for the gap between the simulated ($\sim$23\%)
and measured ($\sim$12\%) recoverable fraction in regime B.

\textbf{(1) Lower $\kappa$ on Ada Lovelace with modern drivers.}
The simulation used $\kappa = 1.5$\,ms, the canonical figure cited in
prior megakernel work for a multi-kernel decode step
\citep{mpk2025,adamk2026}. Measured on the RTX 4050 (Ada Lovelace,
sm\_89) under CUDA 13.3 / driver 610.47, the aggregate overhead of a
48-op sequential chain is $\kappa \approx 0.89$\,ms (regime B) and
$\kappa \approx 0.93$\,ms (regime A) --- roughly $1.6\times$ lower.
Ada Lovelace's improved kernel-dispatch pipeline and the 610-series
driver's reduced host-side submission latency both contribute. Future
measurements on Hopper (H100) and Blackwell-class GPUs may differ.

\textbf{(2) The $\sigma = \kappa/20$ assumption does not hold for CUDA-graph replay.}
The simulation modeled queue-signal cost as $\sigma = \kappa/20$,
reflecting the expectation that a true device-side spin-poll costs
only a queue-write plus a warp wake-up. In the hardware benchmark,
$\sigma$ is measured as the latency of replaying a pre-compiled
CUDA graph of the 48-op chain --- the framework-level analogue of
single-dispatch re-issue. The measured $\sigma/\kappa$ ratio is
$\approx 1/4$ (not $1/20$), because CUDA-graph replay still traverses
the driver's graph-execution path and pays real per-node scheduling
overhead; it is \emph{not} equivalent to device-side residency where
warps never exit the kernel. This is the primary reason the measured
recoverable fraction is roughly half the simulated value.

\textbf{(3) $\sigma$ is a hardware-fixed floor, not workload-dependent.}
Across both regimes, the measured $\sigma$ values are nearly identical:
$\sigma = 0.2213$\,ms (regime A, tool-latency-dominated) and
$\sigma = 0.2204$\,ms (regime B, launch-overhead-dominated). Despite
the two regimes using substantially different tool-latency distributions,
$\sigma$ varies by less than $0.5\%$. This is consistent with CUDA-graph
replay cost being a hardware-and-driver-fixed floor set by the
graph-launch submission path, rather than varying with workload
characteristics. A genuine device-side queue (as in
\texttt{persistent\_kernel.cu}) replaces this floor with the cost of a
single atomic read and a warp-resume signal, which on modern NVIDIA
hardware is measured in single-digit microseconds --- more than an
order of magnitude below the 0.22\,ms floor seen here. The 12.1\%
figure at $T=100$ should therefore be read as a conservative lower
bound on what a full persistent-kernel implementation would achieve.

\textbf{(4) Direct confirmation via device-side instrumentation.}
We instrumented \texttt{persistent\_kernel.cu} with \texttt{clock64()}
timestamps bracketing the queue-poll path --- from the first thread
observing \texttt{SLOT\_READY} on the host-mapped queue, through the
block-wide \texttt{\_\_syncthreads()} barrier, to the start of
\texttt{decode\_segment()} --- and measured the true device-side
$\sigma$ directly, rather than approximating it via CUDA-graph replay.
Across 50 timed steps on the RTX 4050, this gives
$\sigma = 0.095\,\mu\text{s}$ (152 clock cycles/step), roughly
$2{,}300\times$ smaller than the $0.220$\,ms CUDA-graph-replay proxy
used elsewhere in this section. Using this measured $\sigma$ in place
of the proxy raises the recoverable fraction in regime B from
$12.1\%$ to $\mathbf{16.1\%}$ at $T=100$ (with $\kappa$ held at the
synthetic 48-op-chain value from Table~\ref{tab:results}). This
confirms the lower-bound conjecture directly rather than by
extrapolation: true device-side residency recovers meaningfully more
than the CUDA-graph proxy suggested, though the gain here is bounded
by $\kappa$ itself, since $\sigma$ was already a small fraction of
$\kappa$ even under the proxy. A fully consistent measurement --- using
$\kappa$ from the same real forward-pass kernel that produced this
$\sigma$, rather than the synthetic op chain --- reveals an important
distinction, addressed next.

\textbf{(5) Decomposing the saving: fusion vs.\ residency.} We
measured $\kappa$ directly from \texttt{persistent\_kernel.cu}'s own
real forward pass in two configurations: relaunching the
\emph{fused} kernel once per step gives
$\kappa_{\text{fused}} = 1.208\,\mu\text{s}$; keeping that same fused
kernel \emph{resident} (queue-polled, never relaunched) gives the
already-reported $\sigma = 0.095\,\mu\text{s}$. Against the
\emph{unfused} 48-kernel-per-step baseline
($\kappa_{\text{unfused}} = 0.889$\,ms), fusing into one kernel alone
eliminates $99.9\%$ of the overhead
($\kappa_{\text{unfused}} - \kappa_{\text{fused}} = 0.888$\,ms/step)
--- this is precisely the benefit already reported by MPK, Kog, and
Ada-MK \citep{mpk2025,kog2026,adamk2026} for single-turn decode, and
is \emph{not} specific to this paper's proposed mechanism. What
\emph{is} specific to this paper is the further step from
fused-but-relaunched to fused-and-resident:
$\kappa_{\text{fused}} - \sigma = 1.11\,\mu\text{s}$/step, a real but
much smaller saving that, on this hardware and this model size
($N_{\text{LAYERS}}=2$, $\text{DIM}=256$, real compute
$d_t \approx 0.74$\,ms/step), amounts to under $0.1\%$ of a full
decode-plus-tool-call step. The $12$--$16\%$ recovery figures reported
earlier in this section therefore over-attribute credit: they compare
against an \emph{unfused} baseline, which is the regime prior
megakernel work already addresses, rather than isolating the marginal
benefit of residency \emph{beyond} fusion, which is what this paper's
host-queue design specifically contributes. We revise our claim
accordingly below.

\section{Discussion and Limitations}

The design in \S\ref{sec:design} addresses overhead \emph{caused by
relaunching}, not overhead caused by the tool call itself; in the
tool-latency-dominated regime the main win is freeing the GPU for
other tenants during the idle window, not reducing wall-clock latency
for a single agent. An earlier version of the GPU measurement in
Table~\ref{tab:results} timed a single trivial kernel launch
(one \texttt{x.add\_(1.0)} call) as a proxy for $\kappa$, which
underestimated real per-step overhead by approximately 70$\times$
because a realistic decode step dispatches dozens of chained CUDA
kernels --- GEMMs, softmax, layer-norm, residual adds --- not one;
the current table corrects for this by timing a 48-op chain per step,
yielding $\kappa \approx 0.89$--$0.93$\,ms on the RTX 4050, close to the
1--2\,ms range reported in prior megakernel work \citep{mpk2025,adamk2026}
(the gap is attributable to Ada Lovelace's improved dispatch pipeline,
as discussed in \S\ref{sec:discrepancy}). The measured $12$--$16\%$
recovery figures at $T=100$ in regime B compare a fused-and-resident
design against an \emph{unfused}, multi-kernel-per-step baseline, and
so mostly measure the value of \emph{kernel fusion} --- already
established by MPK, Kog, and Ada-MK \citep{mpk2025,kog2026,adamk2026}
--- rather than this paper's specific proposed mechanism. Direct
device-side measurement of \texttt{persistent\_kernel.cu}'s own real
forward pass (\S\ref{sec:discrepancy}, point 5) isolates the marginal
contribution of residency \emph{beyond} fusion: $\kappa_{\text{fused}}
- \sigma = 1.11\,\mu\text{s}$/step, under $0.1\%$ of a full
decode-plus-tool-call step at this model size ($N_{\text{LAYERS}}=2$,
$\text{DIM}=256$). We report this explicitly rather than let the
larger, fusion-driven figure stand as if it were evidence for the
paper's contribution: on current hardware and at this model scale,
persistent residency's benefit \emph{beyond} an already-fused kernel
is real but small, and would need either much smaller $d_t$ (larger,
more heavily fused models where even $\kappa_{\text{fused}}$ is a
larger fraction of step time) or much more frequent step/tool-call
alternation than tested here to become practically significant.
Establishing the regime where this marginal benefit dominates is left
as future work, alongside real forward passes at production model
scale.
The cooperative-yield mechanism (line 6,
Algorithm~\ref{alg:loop}) is the least mature part of the design and
would benefit from integration with existing GPU scheduling
primitives (MPS, MIG, or a custom SM-partition scheme) rather than a
from-scratch implementation. We also do not address correctness under
speculative or partial decode segments that are discarded when a tool
result invalidates an in-flight assumption; this is left as future
work, connected to the broader idea of speculative multi-agent
execution.

\subsection{Broader Applicability: MoE Expert Offload as an Agent Loop}
\label{sec:moe}

The cost model of \S\ref{sec:method} was motivated by tool-calling
agents, but the same structure recurs in a different setting: serving
Mixture-of-Experts (MoE) models on VRAM-constrained hardware via
CPU expert offload (e.g., llama.cpp's \texttt{--n-cpu-moe}). In this
setup, attention and shared layers remain resident on the GPU, but
expert feed-forward weights are kept in system RAM; the router's
per-token, data-dependent choice of which experts to activate
determines which weights must be fetched and computed off-device
before the next layer can proceed. This is structurally the same
alternation as an agent loop --- an on-device segment ($d_t$,
attention/shared compute) followed by an off-device, data-dependent
segment ($\ell_t$, the CPU expert matmul and host round-trip) --- except
that the loop granularity is per-\emph{token}, and per-\emph{layer},
rather than per agent step. Event Tensor \citep{eventtensor2026}
identifies MoE routing as precisely the class of data-dependent
dynamism that static megakernels cannot fuse away, which is the same
obstacle our design encounters at the coarser tool-call granularity.

If this structural analogy holds quantitatively, a persistent-kernel
design analogous to \S\ref{sec:design} --- keeping the GPU-resident
attention/shared-layer kernel warm across each expert round-trip,
rather than relaunching it per layer per token --- would place
CPU-offloaded MoE decoding in our launch-overhead-dominated regime B,
where \S\ref{sec:experiments} measured a lower-bound recovery of
$12.1\%$ at $T=100$ on a synthetic 48-op chain. We flag this
explicitly as an \emph{untested hypothesis}, not a result of this
paper: we have not instrumented a real MoE decode loop (e.g.\ Qwen or
GLM under \texttt{--n-cpu-moe}) to measure whether its per-token
$\kappa,\sigma,d_t,\ell_t$ actually fall in this regime, and the
per-token/per-layer granularity of MoE routing may introduce
overheads (e.g.\ expert-selection variance, cache-locality effects
in CPU RAM) not captured by our agent-step cost model. We leave this
as concrete future work: repeating the measurement methodology of
\S\ref{sec:experiments} on a real MoE inference loop, using the same
reproducibility standard (every reported number backed by a runnable
script) applied throughout this paper.

\section{Conclusion}

Persistent/megakernel designs have closed most of the kernel-launch
overhead gap for single-turn decode. Agent loops introduce a
distinct, largely unaddressed overhead structure at the tool-call
boundary. We formalize this with a simple cost model, propose a
host-queue/resident-kernel design to close the gap, and release a
reference implementation and benchmarking harness so the cost model
can be validated empirically on real hardware rather than asserted.

\bibliographystyle{plainnat}
\begin{thebibliography}{9}

\bibitem[Mirage Persistent Kernel, 2025]{mpk2025}
Mirage Persistent Kernel: A Compiler and Runtime for Mega-Kernelizing
Tensor Programs. \url{https://arxiv.org/html/2512.22219v1}

\bibitem[Kog AI, 2026]{kog2026}
Building a single-kernel, latency-optimized LLM inference engine on
AMD MI300X GPUs. Kog AI Engineering Blog.
\url{https://blog.kog.ai/building-a-single-kernel-latency-optimized-llm-inference-engine-on-amd-mi300x-gpus/}

\bibitem[Ada-MK, 2026]{adamk2026}
Ada-MK: Adaptive MegaKernel Optimization via Automated DAG-based
Search for LLM Inference. \url{https://arxiv.org/html/2605.11581v1}

\bibitem[Event Tensor, 2026]{eventtensor2026}
Event Tensor: A Unified Abstraction for Compiling Dynamic
Megakernels. \url{https://arxiv.org/html/2604.13327v1}

\bibitem[GPUOS, 2026]{gpuos2026}
GPUOS: A GPU Operating System Primitive for Transparent Operation
Fusion. \url{https://arxiv.org/pdf/2604.17861}

\end{thebibliography}

\end{document}
